# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Load data and recode date data to proper format

```{r}
indata<-read.csv("activity.csv")
indata$DT <- strptime(indata$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?
Make a histogram of the total number of steps taken each day. 
Calculate and report the mean and median total number of steps taken per day.
```{r}

agg_bydate<-aggregate(steps ~ date, data=indata, FUN=sum)

hist(agg_bydate$steps)

## Calculate and report the mean and median total number of steps taken per day
mmean<-mean(agg_bydate$steps)
mmedian<-median(agg_bydate$steps)

paste("Mean: ",round(mmean,1))
paste("Median: ",round(mmedian,1))
```

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
agg_byinterval<-aggregate(steps ~ interval, data=indata, FUN=mean)
names(agg_byinterval)[2] <- "MeanSteps"

plot(agg_byinterval$interval,agg_byinterval$MeanSteps, 
     type="l",
     ylab="Average number of steps per interval"
)
```

Let's find which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
paste("Maximum average number of steps is:" ,round(max(agg_byinterval$MeanSteps),1))
paste("at a time interval: ",agg_byinterval[which(agg_byinterval$MeanSteps==max(agg_byinterval$MeanSteps)),1])
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(indata))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in with means. I'll divie an oroginal dataset into two - with NAs and with coplete data. Replace NAs with means for respective time interval. Just copy data for the second data set. Then merge them.
```{r}
m<-merge(indata,agg_byinterval)
mNA<-m[is.na(m$steps), ]
mNNA<-m[!is.na(m$steps), ]
mNA$StepsReplaced<-mNA$MeanSteps
mNNA$StepsReplaced<-mNNA$steps
merged<-rbind(mNNA,mNA)
```
Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. 
```{r}
agg_bydate2<-aggregate(StepsReplaced ~ date, data=merged, FUN=sum)
hist(agg_bydate2$StepsReplaced)

mmean2<-mean(agg_bydate2$StepsReplaced)
mmedian2<-median(agg_bydate2$StepsReplaced)

paste("Mean: ",round(mmean2,1))
paste("Median: ",round(mmedian2,1))
```
Mean is the same, medin become equal to mean


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
merged$WD<-weekdays(merged$DT)
merged$WDD[merged$WD=="Monday"] <- "weekday"
merged$WDD[merged$WD=="Tuesday"] <- "weekday"
merged$WDD[merged$WD=="Wednesday"] <- "weekday"
merged$WDD[merged$WD=="Thursday"] <- "weekday"
merged$WDD[merged$WD=="Friday"] <- "weekday"
merged$WDD[merged$WD=="Saturday"] <- "weekend"
merged$WDD[merged$WD=="Sunday"] <- "weekend"
```
Create timeseries plots for weedays and weekend

```{r}
agg_byintervalWD<-aggregate(StepsReplaced ~ interval, data=merged[merged$WDD=="weekday",], FUN=mean)
agg_byintervalWE<-aggregate(StepsReplaced ~ interval, data=merged[merged$WDD=="weekend",], FUN=mean)

par(mfrow=c(2,1))

plot(agg_byintervalWE$interval,agg_byintervalWE$StepsReplaced, 
     type="l",
     ylab="Number of steps",
     xlab="weekend"
)
plot(agg_byintervalWD$interval,agg_byintervalWD$StepsReplaced, 
     type="l",
     ylab="Number of steps",
     xlab="weekday"
)
```
During weekend object can run during the day. On weekday object run mainly in the morning. 
